---
title: "Data Analysis"
author: "Spiro Stilianoudakis"
date: "9/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(openxlsx)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(sas7bdat)
library(lubridate)
library(tidyr)
library(tableone)
library(ggplot2)
library(grid)
library(gridExtra)
```

# Set working directory

```{r}
setwd("C:/Users/stili/Documents/PAT_project/data")
```

# Reading in data

```{r}
pat_data <- readRDS("C:/Users/stili/Documents/PAT_project/data/pat_data.rds")
```

# Data Analysis

## Is there a difference in sex regarding PAT score

```{r}
ggplot(pat_data, aes(x=Sex, y = PAT, fill=Sex))  + 
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot() +
    geom_signif(test = "t.test", 
                comparisons = list(c("Female", "Male")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    guides(fill=FALSE) +
    theme_minimal()+
    theme_bw() +
    theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

summary(lm(PAT ~ Sex, data=pat_data))

```


## Is there a difference in sex regarding TES score

```{r}
ggplot(pat_data, aes(x=Sex, y = TES, fill=Sex))  + 
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot() +
    geom_signif(test = "wilcox.test", 
                comparisons = list(c("Female", "Male")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    guides(fill=FALSE) +
    theme_minimal()+
    theme_bw() +
    theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

summary(lm(TES ~ Sex, data=pat_data))

```


## Is there a difference in ethnicity regarding PAT score

```{r}
#pat_data$Ethnicity[which(pat_data$Ethnicity=="Black" |
#                           pat_data$Ethnicity=="Other")] <- "Other"
#table(pat_data$Ethnicit)

#pat_data$Ethnicity <- factor(pat_data$Ethnicity, levels=c("White", "Asian", "Other"))

ethnicity_dat <- pat_data[-which(pat_data$Ethnicity=="Unknown"),]
ethnicity_dat$Ethnicity <- factor(ethnicity_dat$Ethnicity, levels=c("White", "African American", "Asian", "Hispanic", "Middle Eastern"))

ggplot(ethnicity_dat, aes(x=Ethnicity, y = PAT, fill=Ethnicity))  + 
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot() +
    geom_signif(test = "t.test", 
                comparisons = list(c("White", "African American"),
                                   c("White", "Asian"),
                                   c("White", "Hispanic"),
                                   c("White", "Middle Eastern"),
                                   c("African American", "Asian"),
                                   c("African American", "Hispanic"),
                                   c("African American", "Middle Eastern"),
                                   c("Asian", "Hispanic"),
                                   c("Asian", "Middle Eastern"),
                                   c("Hispanic", "Middle Eastern")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    guides(fill=FALSE) +
    theme_minimal()+
    theme_bw() +
    theme(axis.text.x = element_text(size=15,
                                     angle = 45,
                                     hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

summary(lm(PAT ~ Ethnicity, data=ethnicity_dat))
```


## Is there a difference in ethnicity regarding TES score

```{r}
ggplot(ethnicity_dat, aes(x=Ethnicity, y = TES, fill=Ethnicity))  + 
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot() +
    geom_signif(test = "t.test", 
                comparisons = list(c("White", "African American"),
                                   c("White", "Asian"),
                                   c("White", "Hispanic"),
                                   c("White", "Middle Eastern"),
                                   c("African American", "Asian"),
                                   c("African American", "Hispanic"),
                                   c("African American", "Middle Eastern"),
                                   c("Asian", "Hispanic"),
                                   c("Asian", "Middle Eastern"),
                                   c("Hispanic", "Middle Eastern")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    guides(fill=FALSE) +
    theme_minimal()+
    theme_bw() +
    theme(axis.text.x = element_text(size=15,
                                     angle = 45,
                                     hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))


summary(lm(TES ~ Ethnicity, data=ethnicity_dat))
```


## Does higher PAT correlate with lower TES (better performance on FM-100)

```{r}
cor(pat_data$TES, pat_data$PAT, method = "spearman")
cor.test(pat_data$TES, pat_data$PAT, method = "spearman")

grob1 = grobTree(textGrob(paste("Spearman Correlation : ", -0.209 ), x = 0.50, y = 0.90, hjust = 0, gp = gpar(col = "red", fontsize = 20)))

grob2 = grobTree(textGrob(paste("p-value : ", "0.0003" ), x = 0.50, y = 0.85, hjust = 0, gp = gpar(col = "red", fontsize = 20)))

ggplot(pat_data,aes(PAT,TES)) + 
  geom_point() +
  #geom_smooth(method='lm',formula=y~x) +
  geom_smooth(method='loess') +
  xlab("PAT Score") +
  ylab("TES Score") +
  theme_minimal()+
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20)) +
  annotation_custom(grob1) +
  annotation_custom(grob2) 
  
summary(lm(TES ~ PAT, data=pat_data))
```


## Is there a difference in age



## Does time of completion correlate with better performance

```{r}
#continuous performance
cor(pat_data$TES, pat_data$Time, method = "spearman")
cor.test(pat_data$TES, pat_data$Time, method = "spearman")

grob1 = grobTree(textGrob(paste("Spearman Correlation : ", -0.279 ), x = 0.50, y = 0.90, hjust = 0, gp = gpar(col = "red", fontsize = 20)))

grob2 = grobTree(textGrob(paste("p-value : ", "< 0.0001" ), x = 0.50, y = 0.85, hjust = 0, gp = gpar(col = "red", fontsize = 20)))

ggplot(pat_data,aes(TES,Time)) + 
  geom_point() +
  #geom_smooth(method='lm',formula=y~x) +
  geom_smooth(method='loess') +
  xlab("TES Score") +
  ylab("Time to completion of FM-100 test") +
  theme_minimal()+
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20)) +
  annotation_custom(grob1) +
  annotation_custom(grob2) 
  
summary(lm(Time ~ TES, data=pat_data))

#categorical performance
ggplot(pat_data, aes(x=Color.Acuity, y = Time, fill=Color.Acuity))  + 
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot() +
    xlab("Color Acuity") +
    guides(fill=FALSE) +
    geom_signif(test = "wilcox.test", 
                comparisons = list(c("Poor", "Average"),
                                   c("Poor", "Superior"),
                                   c("Superior", "Average")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    theme_minimal()+
    theme_bw() +
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

summary(lm(Time ~ Color.Acuity, data=pat_data))
```





